---
title: "SURV727 Final Project Report"
author: "Yael Beshaw"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Github Link 

# Health Communication and Technology in Healthy People Agendas 2020 and 2030
## Abstract
This project aims to analyze the volume of coverage regarding the Health 
Communication and Technology objectives of Healthy People Agendas 2020 and 2030.

# Introduction
Healthy People Agendas for the 2020 and 2030 is a CDC, HHS, and OASH joint
effort to track and achieve health goals for the US Public. Here, we take a look
at a topic relatively new in the scope of public health- Health Communication 
and Technology. In the Healthy People Agendas, objectives that fall under
Health Communication and Technology refer mainly to how the public engages with 
their health using technology and the internet.
## Research Question
How prevalent are these topics across different stakeholders (citizens, news,
lawmakers) and does prevalence differ across the progression of these 
objectives?

# Methods
For the purposes of this project, “coverage” encompasses Google search trends, 
news media coverage, and congressional bills related to these objectives.  
This project utilizes the gtrendsR package to assess results related to keywords
associated with these objectives from Google Trends.  Additionally, it utilizes 
the Dow Jones Factiva API to scrape articles that include the keywords and 
lastly, it observes how many bills were introduced in Congress related to these
topics. We observed 6 objectives, 3 from Healthy People 2020 and 3 from
Healthy People 2030, with one of each in the "Target Met', "No Change", and
"Getting Worse" progression category. 

###Objectives of Interest
#### Healthy People 2020 (January 2015- December 2019)
1. Target Met: 
HC/HIT- 5.1 Persons using the internet to keep track of personal
health information

2. No Change
HC/HIT- 9 
Internet users that can easily access health information.

3. Getting Worse
HC/HIT-6.2 
Persons with broadband Internet access.

#### Healthy People 2030 (January 2020- December 2024)
1. Target Met
HC/HIT- 07
Increase the proportion of adults who use information technology (IT) to track 
health care data or communicate with providers.

2. No Change
HC/HIT- 09 
Increase the proportion of people who can view, download, and send their EHR.

3. Getting Worse
HC/HIT- 05 
Increase the proportion of adults with broadband internet access.

## Data Collection
Data collection was conducting utilizing Google Trends, Dow Jones Factiva, and 
GovTrack.
```{r, echo=FALSE, message=FALSE, error=FALSE}
#load packages
library(readr)
library(tidyverse)
library(gtrendsR)
library(censusapi)
library(dplyr)
library(ggplot2)
library(maps)
library(cowplot)
library(viridis)
library(httr)
library(jsonlite)
```

### Google Trends
Collected data from January 1st 2015 to December 31st 2019 or January 1st 2020
to December 1st 2024 depending on whether the objective is part of 2020 or
2030. For the purposes of this project, two keywords/objective were collected.
#####Healthy People 2020: Target Met
Keywords: “mychart”, “health track”
```{r, echo=FALSE}
Sys.sleep(120)

targetmet_20 <- gtrends(
  c("mychart", "health app"),
  geo="US",
  time= "2015-01-01 2019-12-31",
  low_search_volume = TRUE)
```
#####Healthy People 2020:  No Change
Keywords: "electronic health records", "symptoms"
```{r, echo=FALSE}
Sys.sleep(120)

nochange_20 <- gtrends(
  c("electronic health records", "symptoms"),
  geo="US",
  time= "2015-01-01 2019-12-31",
  low_search_volume = TRUE)
```
#####Healthy People 2020:  Getting Worse
Keywords: “mychart”, “health track”
```{r, echo=FALSE}
Sys.sleep(120)

worse_20 <- gtrends(
  c("broadband internet", "high speed"),
  geo="US",
  time= "2015-01-01 2019-12-31",
  low_search_volume = TRUE)
```
#####Healthy People 2020: Target Met
Keywords: “mychart”, “health track”
```{r, echo=FALSE}
Sys.sleep(120)

targetmet_30 <- gtrends(
  c("telehealth", "health provider"),
  geo="US",
  time= "2020-01-01 2024-12-01",
  low_search_volume = TRUE)
```
#####Healthy People 2020:  No Change
Keywords: "electronic health records", "symptoms"
```{r, echo=FALSE}
Sys.sleep(120)

nochange_30 <- gtrends(
  c("mychart", "health app"),
  geo="US",
  time= "2020-01-01 2024-12-01",
  low_search_volume = TRUE)
```
#####Healthy People 2020:  Getting Worse
Keywords: "broadband internet", "high speed"
```{r, echo=FALSE}
Sys.sleep(120)

worse_30 <- gtrends(
  c("broadband internet", "high speed"),
  geo="US",
  time= "2020-01-01 2024-12-01",
  low_search_volume = TRUE)
```

### Dow Jones 
Collected data from the Dow Jones Factiva for each of the six objectives. For
this project we look at data from the New York Times, Wall Street Journal,
and National Public Radio. Each query was done individually and results
were combined in csv file "hit727".
```{r, echo=FALSE}
hit727 <- read_csv("~/R Scripts and Projects/727_hits.csv", 
                   show_col_types = FALSE)
#dataset can be found in github repository

#Data Collection Method for Dow Jones Factiva
#Queries for each objective were as follows...

#Target 20
#(rst=nytf OR rst=wsjo OR rst=sfnpr) AND (digital near5 health OR manage* near5 
# health OR track* near5 health OR patient near5 portal)

#No Change 20
#(rst=sfnpr OR rst=wsjo OR rst=sfnyt) AND (electronic health records OR EHR OR 
# health literacy OR health education OR accessible health)

#Getting Worse 20
#(rst=sfnpr OR rst=wsjo OR rst=sfnyt) AND (broadband internet OR internet access
# OR digital equity)

#Target 30
# (rst=nytf OR rst=wsjo OR rst=sfnpr) AND (manage* near5 health OR health near5 
# app OR health near5 provider)

#No Change 30
#(rst=nytf OR rst=wsjo OR rst=sfnpr) AND (electronic near5 health near5 records
# OR digital near5 medical records OR accessible near5 health records OR patient 
# near5 portals)

#Getting Worse 30
#(rst=sfnpr OR rst=wsjo OR rst=sfnyt) AND (broadband internet OR internet access
# OR digital equity)
```

### Congressional Terms
Collected data utilizing GovTracker and filter across the past five 
congressional terms (2015-present). Observed 3 subtopics of congressional bills
related to the objectives...
1. Internet, web applications, social media
2. Health information and medical records
3. Health promotion and preventative care 

Congressional Terms
114; 2015-2017
115; 2017-2019
116; 2019-2021
117; 2021-2023
118; 2023-2025

Based on results we create a dataframe containing how many bills were introduced
(House vs Senate), how many were signed into law by the President and which
of the 3 subtopics the bill falls into.
```{r, echo=FALSE}
bills727 <- read_csv("~/R Scripts and Projects/727_bills.csv",
                     show_col_types = FALSE)
#dataset can be found in github repository
```

## Analysis and Results
Compares the volume of coverage for objectives between both agendas and across
their respective progression status.

#### Google Trends
This plot reflects the mean number of hits across the respective 5 year periods
of data I have for each agenda. We see that across the board there is relatively
similar interest with broadband internet topics being high despite their 
lackluster progression. 
```{r, message=FALSE, error=FALSE, warning=FALSE}

#create function that twill allows us to pull out the total hits/ query
df2tibble <- function(df) {
  tibble_data <- as_tibble(df$interest_over_time) %>%
    drop_na() %>%
    mutate(hits = as.numeric(hits)) %>% 
    mutate(hits = replace_na(hits, 0))
  return(tibble_data)
}
targetmet_20_time <- df2tibble(targetmet_20)
nochange_20_time <- df2tibble(nochange_20)
worse_20_time <- df2tibble(worse_20)
targetmet_30_time <- df2tibble(targetmet_30)
nochange_30_time <- df2tibble(nochange_30)
worse_30_time <- df2tibble(worse_30)

#create a function that will give us mean,median, and variance for each
descriptive_hits <- function(df) {
  df%>%
    summarize(
      mean= mean(df$hits),
      median= median(df$hits),
      variance= var(df$hits)
  )
}

#bind the results into a dataframe
stats <- bind_rows(
  descriptive_hits(targetmet_20_time) %>% mutate(dataset = "2020 Target Met"),
  descriptive_hits(nochange_20_time) %>% mutate(dataset = "2020 No Change"),
  descriptive_hits(worse_20_time) %>% mutate(dataset = "2020 Worse"),
  descriptive_hits(targetmet_30_time) %>% mutate(dataset = "2030 Target Met"),
  descriptive_hits(nochange_30_time) %>% mutate(dataset = "2030 No Change"),
  descriptive_hits(worse_30_time) %>% mutate(dataset = "2030 Worse")
)

library(ggplot2)

ggplot(stats, aes(x = factor(dataset, levels = c("2020 Target Met", 
                                                 "2020 No Change",  
                                                 "2020 Worse",  
                                                 "2030 Target Met",  
                                                 "2030 No Change",  
                                                 "2030 Worse")),
                  y = mean, fill = dataset)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(mean, 1)),
            position = position_dodge(width = 0.9), vjust = -0.5, color = "black", size = 3) +
  labs(
    title = "Mean Number of Hits Across Time for Google Trends",
    x = "objective status",
    y = "mean hits",
    fill= "objective"
  ) +
  scale_fill_manual(values = c(
    "2020 Target Met" = "darkgreen",
    "2020 No Change" = "gold",
    "2020 Worse" = "darkred",
    "2030 Target Met" = "darkgreen",
    "2030 No Change" = "gold",
    "2030 Worse" = "darkred"
  )) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

```

Here, we assess the total number of hits and mapped it across the states to
observe where the Google trends data volume is coming from. We do this in order
to salience of these topics within the general public and locate where in the
general public we observe this.We see that for target met objectives there is 
a spread of states with 120 hits and above in both agenda periods. For no 
progress we see there is a belt of high hits from northeast to south in 
2015-2020 but then it seemingly rotates to density in Wisconsin to South
Carolina in 2020-2024. Lastly, for worsening progress, we see that the searches
are dense across the board and in states such as California and Oregon which 
are not usually associated with the "rural and poor" label that comes with a 
lack of internet. 
```{r}
#create a function that will allow us to pull out hits/state(region)
df2tibble_state <- function(df) {
  tibble_data <- as_tibble(df$interest_by_region) %>%
    drop_na() %>%
    mutate(hits = as.numeric(hits)) %>% 
    mutate(hits = replace_na(hits, 0))
  return(tibble_data)
}

targetmet_20_state <- df2tibble_state(targetmet_20)
nochange_20_state <- df2tibble_state(nochange_20)
worse_20_state <- df2tibble_state(worse_20)
targetmet_30_state <- df2tibble_state(targetmet_30)
nochange_30_state <- df2tibble_state(nochange_30)
worse_30_state <- df2tibble_state(worse_30)

#create a function that takes the total number of hits and organizes it
#from greatest to least 
state_hits <- function(df) {
  df%>%
    group_by(location) %>%
    summarise(total_hits = sum(hits, na.rm = TRUE)) %>%
    arrange(desc(total_hits))
}

targetmet_20_state <- state_hits(targetmet_20_state)
nochange_20_state <- state_hits(nochange_20_state)
worse_20_state <- state_hits(worse_20_state)
targetmet_30_state <- state_hits(targetmet_30_state)
nochange_30_state <- state_hits(nochange_30_state)
worse_30_state <- state_hits(worse_30_state)

#map the data onto the US map
us_states <- map_data("state")

#target20
data_target20 <- targetmet_20_state %>%
  mutate(location = tolower(location))

map_data <- us_states %>%
  rename(region = region) %>%
  left_join(data_target20, by = c("region" = "location"))

target20_map <- ggplot(map_data, aes(x = long, y = lat, group = group, fill = total_hits)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(
    low = "lightblue",
    high = "darkblue",
    na.value = "gray90"
  ) +
  labs(title = "Total Hits by State", fill = "Total Hits") +
  theme_minimal()


#nochange20
data_nochange20 <- nochange_20_state %>%
  mutate(location = tolower(location))

map_data <- us_states %>%
  rename(region = region) %>%
  left_join(data_nochange20, by = c("region" = "location"))

nochange20_map <-ggplot(map_data, aes(x = long, y = lat, group = group, fill = total_hits)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(
    low = "lightyellow", 
    high = "darkorange",
    na.value = "gray90"
  ) +
  labs(title = "Total Hits by State", fill = "Total Hits") +
  theme_minimal()


#worse20
data_worse20 <- worse_20_state %>%
  mutate(location = tolower(location))

map_data <- us_states %>%
  rename(region = region) %>%
  left_join(data_worse20, by = c("region" = "location"))

worse20_map <- ggplot(map_data, aes(x = long, y = lat, group = group, fill = total_hits)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(
    low = "lightpink", 
    high = "darkred",
    na.value = "gray90"
  ) +
  labs(title = "Total Hits by State", fill = "Total Hits") +
  theme_minimal()


#target30
data_target30 <- targetmet_30_state %>%
  mutate(location = tolower(location))

map_data <- us_states %>%
  rename(region = region) %>%
  left_join(data_target30, by = c("region" = "location"))

target30_map<- ggplot(map_data, aes(x = long, y = lat, group = group, fill = total_hits)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(
    low = "lightblue",
    high = "darkblue",
    na.value = "gray90"
  ) +
  labs(title = "Total Hits by State", fill = "Total Hits") +
  theme_minimal()


#nochange30
data_nochange30 <- nochange_30_state %>%
  mutate(location = tolower(location))

map_data <- us_states %>%
  rename(region = region) %>%
  left_join(data_nochange30, by = c("region" = "location"))

nochange30_map<- ggplot(map_data, aes(x = long, y = lat, group = group, fill = total_hits)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(
    low = "lightyellow", 
    high = "darkorange",
    na.value = "gray90"
  ) +
  labs(title = "Total Hits by State", fill = "Total Hits") +
  theme_minimal()


#worse30
data_worse30 <- worse_30_state %>%
  mutate(location = tolower(location))

map_data <- us_states %>%
  rename(region = region) %>%
  left_join(data_worse30, by = c("region" = "location"))

worse30_map <-ggplot(map_data, aes(x = long, y = lat, group = group, fill = total_hits)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(
    low = "lightpink", 
    high = "darkred",
    na.value = "gray90"
  ) +
  labs(title = "Total Hits by State", fill = "Total Hits") +
  theme_minimal()

combined_plot <- plot_grid(
  target20_map, target30_map,
  nochange20_map, nochange30_map,
  worse20_map , worse30_map,
  ncol = 2,
  align = "hv",
  label_size = 8
)

combined_plot
```

#### Dow Jones
We see again that there is high interest for broadband internet topics but 
surprising much lower interest for our in progress topic of accessing health 
information (EHR) by assessing the mean number of articles per objective.
```{r}
#gather the values for the mean number of articles per objective
mean_hits <- hit727 %>%
  group_by(objective) %>%
  summarize(mean = mean(hits))%>%
  mutate(objective = factor(objective, 
                            levels = c("target20", "nochange20", "worse20", 
                                       "target30", "nochange30", "worse30"), 
                            labels = c("2020 Target Met", "2020 No Change",
                                       "2020 Worse", 
                                       "2030 Target Met", "2030 No Change",
                                       "2030 Worse")))
#plot
ggplot(mean_hits, aes(x = factor(objective, levels = c("2020 Target Met", 
                                                 "2020 No Change",  
                                                 "2020 Worse",  
                                                 "2030 Target Met",  
                                                 "2030 No Change",  
                                                 "2030 Worse")),
                  y = mean, fill= objective)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(mean, 1)),
            position = position_dodge(width = 0.9), vjust = -0.5, 
            color = "black", size = 3) +
  labs(
    title = "Mean Number of Articles Across Time",
    x = "objective status",
    y = "mean"
  ) +
  scale_fill_manual(values = c(
    "2020 Target Met" = "darkgreen",
    "2020 No Change" = "gold",
    "2020 Worse" = "darkred",
    "2030 Target Met" = "darkgreen",
    "2030 No Change" = "gold",
    "2030 Worse" = "darkred"
  )) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")
```
We then look at this plot below to observe the overall trend of these
objectives in news media. As a whole, we see a decrease in articles over time, 
but we see that this decrease occurs at different rates per objective.
```{r}
#observe trends for number of articles over time
ggplot(hit727, aes(x = date, y = hits, color = objective, group = objective)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE, aes(group = objective), linetype = "dotted") +
  labs(title = "Articles Over Time by Objective", x = "year", y = "article hits") +
  scale_color_manual(values = c(
    "target20" = "blue", 
    "nochange20" = "orange", 
    "worse20" = "red", 
    "target30" = "deepskyblue3", 
    "nochange30" = "darkorange", 
    "worse30" = "deeppink"
  )) +
  scale_x_continuous(breaks = seq(min(hit727$date), max(hit727$date), by = 1)) +
  theme_minimal() 
```
### Congressional Terms

Here, we observe the total number of bills introduced by the House and Senate
across each topic. We compare that to the percentage of bills that are
enacted (signed by the President) across topic and congressional term. We 
see here that bills regarding the internet are much less than that of 
medical information or health promotion. Despite this we still see a simillar
percentage of laws that are enacted across topic. However, since the 117th
congressional term, the percentage of bills enacted regarding the internet
have been much lower.
```{r}
#create a column that assesses the percentage of bills enacted for 
#each topic/ term
bills_percent <- bills727 %>%
  mutate(total= (house_count + senate_count))

bills_percent <- bills727 %>%
  mutate(total= (house_count + senate_count)) %>%
  mutate(percent_enacted = (enacted_pres / total) * 100)

#plot
colors_custom <- c(
  "infomed" = "lightblue",
  "healthpromo" = "orange",
  "internet" = "lightpink"
)

percent_enacted<- ggplot(bills_percent, aes(x = factor(congress), y = percent_enacted, fill = subject)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values=colors_custom)+
  labs(
    title = "Percent Enacted",
    x = "Term",
    y = "Percentage Enacted (Signed by President)",
    fill = "Topic"
  ) +
  theme_minimal()

total<- ggplot(bills_percent, aes(x = subject, y = total, fill = subject)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values=colors_custom)+
  labs(
    title = "Total Bills across Subject",
    x = "Topic",
    y = "Total Number of Bills",
    fill = "Topic"
  ) +
  theme_minimal()+
  theme(legend.position = "none")

combined <- plot_grid(
  total, percent_enacted,
  ncol = 2,
  align = "hv",
  label_size = 8
)
combined
```
#Conclusions
Based on these findings, we are able to assert that objectives across
progression status have similar prevalence across the stakeholders we observed
(general public, news media, and lawmakers). Thus, we can not certainly say that
an objective that has worsening progress is not an objective that is salient.
Instead, these results show us that there is an interest regarding objectives
that are not progressing well and that there is an opportunity to use that
interest to assist with progression efforts. For example, as seen in the 
US Map, interest regarding broadband internet is across the nation and in very
diverse states. We may be able to focus efforts in states that are already
very interested in this topic to address progress and expand methods outward.
Additionally, we see that news media does not favor the objective that has 
no change/progress, as the mean number of articles published relating to it
is much smaller than that of the other two. This indicates that we have an 
opportunity to encourage news media to talk more about topics relating to this
subject, as searches related to it are still high based on Google Trends 
results. We hope that increased news coverage may be able to address the 
interest and concerns that the public has and thus influence lawmakers to do 
more regarding these topics.

#References
https://www.cdc.gov/nchs/healthy_people/hp2020/progress-tables.htm 

https://odphp.health.gov/healthypeople/objectives-and-data/browse-objectives/health-communication

https://global-factiva-com.proxy-um.researchport.umd.edu/sb/default.aspx?NAPC=S  

https://www.govtrack.us/ 
